---
title: "Ruth Holley-Assignment 3"
author: "Ruth Holley"
date: "5/9/2019"
output: html_document
---
In this homework, you will do some data analysis using R
for the Forest Fire Data described
https://archive.ics.uci.edu/ml/datasets/forest+fires
The dataset is used to find the relationship between the
burned area of forest fires and meteorological data. 

1. Import the data into R.
```{r }
library(dplyr)

# loading dataset
Forest = read.csv("/Users/candyland90/Downloads/forestfires.csv", header = TRUE, sep = ',')
Forest

```

a. How many observations are there in the dataset?
```{r cars}
str(Forest)
dim(Forest)
#There are 517 observations in this dataset
```

b. How many observations are there with a fire (i.e., area>0)
```{r}
forestfires <- Forest %>% filter(area > 0) 
str(forestfires)
#There are 270 observations with a fire
```

c. How many observations are there with a rain (i.e., rain>0)
```{r}
forestrain <- Forest %>% filter(rain > 0) 
str(forestrain)
#There are 8 observations with rain
```

d. How many observations are there with both a fire and a rain? 
```{r }
forestfires_rain <- Forest %>% filter(rain > 0 , area > 0) 

str(forestfires_rain)

#There are 2 observations that are both fire and rain
```

2. Show the columns month, day, area of the all the observations. 
```{r}
Forest[, c('month', 'day', 'area')]

Forest %>% select(month, day, area)
```

3. Show the columns month, day, area of the observations with a fire.
```{r}
Forest[Forest$area>0, c('month', 'day', 'area')]

Forest %>% filter(area>0) %>% select(month, day, area)
```

4. How large are the five largest fires (i.e., having largest area)
```{r}
Forest %>% arrange(desc(area)) %>% top_n(5) %>% select(area)
```

a. What are the corresponding month, temp, RH, wind, rain area?
```{r}
Forest %>% arrange(desc(area)) %>% top_n(5) %>% select(month, temp, RH, wind, rain, area)
```

b. Add one column to the data indicating whether a fire occurred for each observation (True for area > 0 and False for area ==0) 
```{r}
Forest$fire = factor(Forest$area>0)

Forest = Forest %>% mutate(fire=area>0)
str(Forest)

Forest = Forest %>% mutate(fire=area==0)
Forest$fire = factor(Forest$fire)
str(Forest)
```
5. Construct a matrix with 10 columns and 10 rows, filled with random numbers between 0 and 1. Apply two different methods.
#Method 1
```{r}
m <- matrix(runif(100), nrow=10, ncol=10)
m
```

#Method 2
```{r}
m1 <-matrix(sample(0:1, 100, replace=TRUE), 10, 10)
m1
```
a. Calculate the row means of this matrix.
```{r}
#Using the method 1 matrix from above.
rowMeans(m)
```
b. Remake the above matrix, with 100 columns, and 10 rows. Then calculate the column means.
```{r}
m <- matrix(runif(1000), nrow=10, ncol=100)
m
```

```{r}
colMeans(m)
```

