---
title: "Assignment 4"
author: "Ruth Holley"
date: "5/27/2019"
output: html_document
---
Load worldrecord.csv and answer the following?
```{r }
library(dplyr)

worldrecord <- read.csv("/home/ruth/Downloads/WorldRecords.csv",header = TRUE, sep = ',')

summary(worldrecord)
str(worldrecord)
dim(worldrecord)
```
1.How many different types of events (e.g.“Mens 100m”,	 “Womens	shotput” etc) are represented	in the dataset.

There are 10 different events represented in the dataset.
```{r }
table(worldrecord$Event)
```

2. In what year did Usain Bolt first break the	world	record	for	the	Men’s	100m?
2008
```{r}
#select(worldrecord$Athlete =="Usain Bolt")
 filter(worldrecord,Athlete =="Usain Bolt", Event=="Mens 100m",Record < 9.74)
```

3. Which variable tells us the record	setting	time or	distance?	The	variable name	in the data	set	is?	What type	of	the	variable is	this?

Record and it is a numeric
```{r}
#Record
class(worldrecord$Record)
```
4.Create a subset of	the	dataset	that contains only the	world	record cases for men’s shotput and women’s shotput
```{r}
filter(worldrecord,Event %in% c("Mens Shotput", "Womens Shotput"))
```

5. Create	a	scatter plot of the	year and record shotput	distance one for men and one for women.
```{r}
#For Men
#Subset the data
menshot <- worldrecord[worldrecord$Event=='Mens Shotput', ]

plot(menshot$Year,menshot$Record,main='Mens Shotput World Records',xlab='Year',ylab='Record Shotput Distance')
```

```{r}
#For Women
#Subset the data
womenshot <- worldrecord[worldrecord$Event=='Womens Shotput', ] 

plot(womenshot$Year,womenshot$Record,main='Womens Shotput World Records',xlab='Year',ylab='Record Shotput Distance')
```

6. Find the average/mean time	for each event.	How many	 athletes have time more than average in	each event.

#Step 1: Find the average/mean time for each event
```{r}
worldrecord%>%
select(Event,Type,Record,Athlete)%>%
group_by(Event)%>%
filter(Type =="time")%>%
summarize(average=mean(Record))
```

#Step 2: How many	athletes have time more than average in	each event. 
```{r}
b1<-worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  group_by(Event)%>%
  filter(Event == "Mens 100m" & Record >= 9.848824)%>%
  summarise(Gt_Than_Avg=n())
b1
#There are 9 athletes that have time greather than the average for this event
```

#There are 10 athletes that have time greather than the average for this event(b2)
```{r}
b2<-worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  group_by(Event)%>%
  filter(Event == "Mens 800m" & Record >= 105.345833)%>%
  summarise(Gt_Than_Avg=n())
b2
```

#There are 15 athletes that have time greather than the average for this event(b3)
```{r}
b3<-worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  group_by(Event)%>%
  filter(Event == "Mens Mile" & Record >= 237.493750)%>%
  summarise(Gt_Than_Avg=n())
b3
```

#There are 6 athletes that have time greather than the average for this event(b4)
```{r}
b4<-worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  group_by(Event)%>%
  filter(Event == "Womens 100m" & Record >= 10.880000)%>%
  summarise(Gt_Than_Avg=n())
b4
```

#There are 13 athletes that have time greather than the average for this event(b5)
```{r}
b5<-worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  group_by(Event)%>%
  filter(Event == "Womens 800m" & Record >= 127.851724)%>%
  summarise(Gt_Than_Avg=n())
b5
```
#There are 5 athletes that have time greather than the average for this event(b6)
```{r}
b6<-worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  group_by(Event)%>%
  filter(Event == "Womens Mile" & Record >= 263.653846)%>%
  summarise(Gt_Than_Avg=n())
b6
```

7. Select	the	athlete	who	took most	time in men’s	100m and women’s event.
#Bob Hayes took the most time in Mens 100m with a record time of 10.06
```{r}
#Mens
worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  filter(Event == "Mens 100m")
```
#Womens
There is a tie on who took the most time in the timed
women's events. Wyomia Tyus and Renate Stecher both have time of 11.07
```{r}
worldrecord%>%
  select(Event,Type,Record,Athlete)%>%
  filter(Event %in% c("Womens 100m", "Womens 100m", "Womens Mile"))
```

8. Which country won maximum times of	men’s 100m event?
The United States with 11 wins.
```{r}
worldrecord%>%
  select(Event,Record,Nationality,Athlete)%>%
group_by(Nationality)%>%
filter(Event == "Mens 100m")%>%
summarise(Country_wins=n())
```

9. How many athletes are there in	each event?
```{r}
worldrecord%>%
  select(Event,Athlete)%>%
  group_by(Event)%>%
summarise(Total_Athlete=n_distinct(Athlete))
```

10. Which	country	has	maximum	wins?
#URS has maximum wins with a total of 49
```{r}
worldrecord%>%
  select(Event,Record,Nationality,Athlete,Year)%>%
  group_by(Nationality)%>%
summarise(max_wins=n())
```

