---
title: "Assignment 2"
author: "Ruth Holley"
date: "4/30/2019"
output: html_document
---

1. Write a	program	to find	the	outliers using Box	and	Whisker’s	criterion	discussed	in the class.
```{r }
data1 <- c(22,30,35,40,40,52,58,60,65,72,250)
summary(data1)
boxplot(data1)
```

2. Make	two	vectors: x <-c(1,2,5,10,12)  y <-c(2,5,1,0,12)
```{r }
x <-c(1,2,5,10,12)
y <-c(2,5,1,0,12)

#A. Find the	values that	are	contained	in both	X	and	Y.
intersect(x,y)
#B. Find values that are in x but not	in y and (vice versa)
union(setdiff(x,y), setdiff(y,x))
```
#C. Construct	a	vector	that	contains	all	values	contained	in	either X	or Y, and	compare	this	vector	to c(X,Y)
```{r}
a <- c(intersect(x,y),setdiff(x,y),setdiff(y,x))
a
b <- c(x,y)
b
a == b
unique(b[! b %in% a])
```

Load USArrests data	set.	
```{r }
data(USArrests)
USArrests
```
3. Which states	has	most and least assault, murder, and rape arrests.
```{r}
# States with most Assault, murder, rape
mAssault <- which(USArrests$Assault == max(USArrests$Assault))
rownames(USArrests)[mAssault]

mMurder<- which(USArrests$Murder == max(USArrests$Murder))
rownames(USArrests)[mMurder]

mRape <- which(USArrests$Rape == max(USArrests$Rape))
rownames(USArrests)[mRape]
```

#States with the least Assault, murder, rape

```{r}
lAssault <- which(USArrests$Assault == min(USArrests$Assault))
rownames(USArrests)[lAssault]

lMurder<- which(USArrests$Murder == min(USArrests$Murder))
rownames(USArrests)[lMurder]

lRape <- which(USArrests$Rape == min(USArrests$Rape))
rownames(USArrests)[lRape]
```

4. Which states	are	in the bottom	25%	of murder	
```{r}
summary(USArrests)
#Bottom 25% of murder
bottomMurderRt <- quantile(USArrests$Murder)[2]
bottomMurderRt
subset(USArrests, Murder < bottomMurderRt, select = Murder)
```

and in the top 25% of murder.
```{r}
#Top 25% of murder
topMurderRt <- quantile(USArrests$Murder)[4]
topMurderRt
subset(USArrests, Murder > topMurderRt, select = Murder)
```

5. The following	function	calculates the	mean	and	standard	deviation	of a numeric vector

Fn1<-function(x){
Mean1<- mean(x)
Sd1<-sd(x)
return(mean	=	mean1,	sd	=	sd1)
}
Modify	the	function	so	that:	(a)	the default	is	to	use	rnorm to	generate	30 random	normal	numbers,	and	return	the	standard	deviation.
```{r}
Fn1 <- function(x){
  x = rnorm(30)
  sd1 <- sd(x)
  return(sd = sd1)
}
```

(b)	if	there	are	missing	values,	the	mean	and	standard	deviation	are	calculated	for	the	remaining	values.
```{r}
Fn2 <- function(x){
  x = is.na(rnorm(30))
  mean1 <- mean(x)
  sd1 <- sd(X)
  return(mean = mean1, sd = sd1)
}
```


